---
title: "Rosters and Units"
author: "Roman Rivera"
date: "1/20/2018"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 7
    fig_height: 4
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include=TRUE, warning = F, message = F)
library(tidyverse)
library(magrittr)
library(lubridate)
library(ggthemes)
theme_set(theme_igray())
library(knitr)
library(kableExtra)
library(gridExtra)
library(xts)
library(timetk)
library(dygraphs)
source('../helper.functions.R')
```

# Introduction
```{r introduction, cache=TRUE}
prof <- read_csv("../../full-output/profiles/final-profiles.csv.gz") %>%
        mutate(gender = recode(gender, 
                               MALE = 'Male', FEMALE = 'Female', 
                               .default='Unknown'),
               race = recode(race, 
                             BLACK = 'Black', HISPANIC = 'Hispanic', WHITE = 'White', 
                             .default = 'Other'),
               resignation_date = ymd(resignation_date),
               appointed_date = ymd(appointed_date))
```
In the data we have collected, there are `r nrow(prof)` officers whose profiles have been aggregated across, at most, `r max(prof$profile_count)` files. These officers were appointed between `r niceDate(min(prof$appointed_date, na.rm=T), 'ymd')` and `r niceDate(max(prof$appointed_date, na.rm=T), 'ymd')`, with `r pnum(nrow(subset(prof, is.na(appointed_date))))` officers lacking appointed date information. Additionally, there are known resignation dates for `r  pnum(nrow(subset(prof, !is.na(resignation_date))))` (`r  pcnt(nrow(subset(prof, !is.na(resignation_date)))/nrow(prof))`%) of officers, and imputed current (most recently known) units for `r pnum(nrow(subset(prof, !is.na(current_unit))))` officers (`r  pcnt(nrow(subset(prof, !is.na(current_unit)))/nrow(prof))`%). Race and gender information is known for `r pnum(nrow(subset(prof, !is.na(race))))` (`r  pcnt(nrow(subset(prof, !is.na(race)))/nrow(prof))`%) and `r pnum(nrow(subset(prof, !is.na(gender))))` (`r  pcnt(nrow(subset(prof, !is.na(gender)))/nrow(prof))`%), respectively. Birth year information is known for `r pnum(nrow(subset(prof, !is.na(birth_year))))` (`r  pcnt(nrow(subset(prof, !is.na(birth_year)))/nrow(prof))`%) officers. Full demographic (race, gender, birth year, appointed date) information exists for `r pnum(nrow(subset(prof, !is.na(race) & !is.na(gender) & !is.na(birth_year) & !is.na(appointed_date))))` (`r  pcnt(nrow(subset(prof, !is.na(race) & !is.na(gender) & !is.na(birth_year) & !is.na(appointed_date)))/nrow(prof))`%) of officers.

# Composition
The CPD has seen significant demographic shifts over the last 50 years. Looking at the racial and gender composition of officers based on the decade of their appointment...
```{r appointed_decades, cache=TRUE, dependson="introduction"}
prof %>% 
  filter(! (is.na(appointed_date) | is.na(race) | is.na(gender))) %>% 
  mutate(appointed_decade = ymd(paste0((year(appointed_date) %/% 10 * 10), '-01-01')),
         rxg = str_wrap(str_to_title(paste0(race, ' x ', gender)), width=30)) %>%
  ggplot(., aes(x = appointed_decade, fill = rxg)) + 
  geom_bar(position='fill') + 
  scale_y_continuous(labels=scales::percent) + 
  scale_fill_brewer(palette="Paired") +
  theme(legend.position = 'bottom') +
  labs(x = 'Decade of Appointment', y = '% of Officers Appointed', fill='Race and Gender')
```

```{r active_officers_overtime, cache=TRUE, dependson="introduction"}
proft <- prof %>% 
  mutate(resignation_date = if_else(!is.na(resignation_date), 
                                    ymd(resignation_date), 
                                    ymd('2017-12-31')),
         start_year = year(appointed_date),
         end_year = year(resignation_date)) %>%
  filter(! is.na(start_year), 
         start_year <= end_year, 
         end_year < start_year + 50,
         ! is.na(race),
         ! is.na(gender)) %>%
  select(UID, race, gender, start_year, end_year) %>%
  mutate(active_dates = map2(start_year, end_year, seq)) %>% 
  unnest() %>% 
  mutate(active_dates = ymd(paste0(active_dates, '-01-01'))) %>%
  mutate(rxg = str_to_title(paste0(race, ' x ', gender)))

proft %>% 
  ggplot(., aes(x=active_dates, fill=rxg)) + 
  geom_bar() +
  scale_fill_brewer(palette="Paired") +
  theme(legend.position = 'bottom') +
  labs(x = 'Year', y='Number of Officers Active', fill='Race and Gender')
```
```{r active_officers_dygraph, cache=TRUE, dependson="active_officers_overtime"}
proft %>%
  group_by(active_dates, race) %>% tally() %>%
  select(Year = active_dates, race, n) %>%
  spread(race, n)  %>% 
  replace_na(list("Black" = 0, "Hispanic"=0, "Other"=0, "White"=0)) %>% 
  tk_xts(.) %>% 
  dygraph() %>%
  dySeries("Black", label = "Black") %>%
  dySeries("White", label = "White") %>%
  dySeries("Hispanic", label = "Hispanic") %>%
  dySeries("Other", label = "Other") %>%
  dyLegend(width = 500) %>%
  dyRangeSelector(dateWindow = c(min(proft$active_dates), max(proft$active_dates)))
```

# Recent Officers
```{r recent_officers, cache=TRUE, dependson="introduction"}
min_date <- ymd('2005-01-01')
max_date <- ymd('2016-12-31')
profs <- prof %>% 
      filter((is.na(resignation_date) | resignation_date >= min_date),
             (! is.na(appointed_date) & appointed_date <= max_date),
             ! is.na(race),
             ! is.na(gender),
             ! is.na(birth_year)) %>%
      mutate(tenure = if_else(is.na(resignation_date) | resignation_date <= appointed_date, 
                              interval(ymd(appointed_date), ymd(max_date))/years(1),
                              interval(ymd(appointed_date), ymd(resignation_date))/years(1)),
             age_at_appointment = year(appointed_date) - birth_year) %>%
      filter(tenure < 45,
             between(age_at_appointment,20, 50)) %>%
      mutate(`Tenure` = cut(tenure, 
                                breaks = c(0,5, 10, 15, 20, 45), 
                                include.lowest=TRUE),
             `Age at Appointment` = cut(age_at_appointment, 
                                breaks = c(20,25, 30, 40, 50), 
                                include.lowest=TRUE))

rxg_table <- pcnt_table(profs$race, profs$gender, prop_margin = 1)
aap_table <- profs %>% group_by(race) %>% summarise(aap = round(mean(age_at_appointment), 2))
tenure_table <- profs %>% group_by(race) %>% summarise(tenure = round(mean(tenure), 2)) %>% arrange(-tenure)
```
Looking at the `r pnum(nrow(profs))` officers active from  `r niceDate(min_date, 'ymd')` to `r niceDate(max_date,'ymd')`, with non-missing demographic information and reasonable tenures and ages at appointment. These officers are roughly `r tableToFacts(pcnt_table(profs$race), in_parans=F)`. Women are significantly underrepresented in the CPD, only making up `r  tableToFacts(pcnt_table(profs$gender), c('Female'), in_parans=F)` of all officers, though the gender imbalance is least extreme in Black officers, being `r rxg_table['Black', 'Female']`% female, compared to White and Hispanic officers, `r rxg_table['White', 'Female']`% and `r rxg_table['Hispanic', 'Female']`% female, respectively. `r subset(aap_table, aap == max(aap))$race` officers tend to be the oldest when appointed (`r max(aap_table$aap)`) years old on average), while (`r subset(aap_table, aap == min(aap))$race` officers tend to be the youngest (`r min(aap_table$aap)` years old on average). `r tenure_table[1,'race']` officers are the most tenured on average at `r tenure_table[1,'tenure']` years, followed by `r tenure_table[2,'race']` (`r tenure_table[2,'tenure']` years), then `r tenure_table[3,'race']` and `r tenure_table[4,'race']` at `r tenure_table[3,'tenure']` and `r tenure_table[4,'tenure']` years, respectively.

```{r recent_officers_table, cache=TRUE, dependson="recent_officers", results='asis'}
niceTableStacked(profs, 'race', c("gender", "Age at Appointment", "Tenure"), scroll=F)
```